---
title: "typewriter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{typewriter}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(typewriter)
```

The goal of typewriter is to add type safety to your R code. Under the hood, it mainly uses base R functions to keep its dependencies as low as possible. It lets you

* add types checks to your function arguments with `check_args()`
* create typed structures with `typed_struct()` and `base_model()`

### Typed structures

To make it adding type safety as fun as possible, a typed structure is just an extended R list where types are defined as _type check functions_ (like `is.integer`) or _type strings_ (like `"integer"`):

```{r}
my_type <- typed_struct(
  a_number = "integer",
  some_text = "character"
)


obj <- my_type(
  a_number = 1L,
  some_text = "typewriter"
)

obj

class(obj)
```

Try to assign a character value to an integer field:

```{r}
try(
  obj$a_number <- "Hi"
)
```

Try to create a new object with an invalid type:

```{r}
try(
  my_type(
    a_number = 1,
    some_text = "typewriter"
  )
)
```

If you use _type strings_ as above, a `typed_struct` is very close to structures in type safe programming languages like _Rust_ or _Go_:

```rust
# Rust
struct Person {
    name: String,
    age: u8
}
```

```go
# Go
type Person struct {
  name string
  age int
}
```

```{r}
# R
Person <- typed_struct(
  name = "character",
  age = "integer"
)

# R, alternate syntax
Person <- typed_struct(
  name = character(),
  age = integer()
)
```

A typed structure could be used as input for a function:

```{r}
say_hello_to <- function(person) {
  stopifnot(inherits(person, "Person"))
  paste("Hello", person$name, "you are", person$age, "years old!")
}

hanna <- Person(name = "Hanna", age = 10L)
say_hello_to(hanna)

try(say_hello_to(list(name = "Peter", age = 20L)))
```

You can also use it to validate a data frame:

```{r}
df <- data.frame(
  name = c("Peter", "Hanna"),
  age = c(12L, 10L)
)

Person(df)

df$id <- 1:2
try(Person(df))
```

A type can also be defined as a function that takes the value to be validated as its only argument and returns `TRUE` in case of success:

```{r}
my_type <- typed_struct(
  a_number = is.integer,
  some_text = is.character,
  # Add a range check to y
  y = function(y) {
    is.integer(y) & y > 1 & y < 10
  }
)

obj <- my_type(
  a_number = 1:2,
  some_text = "typewriter",
  y = 5L
)

obj

try(
  my_type(
    a_number = 1L,
    some_text = "typewriter",
    y = 1L
  )
)
```

### Typed functions

With `check_args()` you can add type checks to your functions:

```{r}
add_two_numbers <- function(a, b) {
  check_args(
    a = "integer",
    b = "integer"
  )
  a + b
}

add_two_numbers(10L, 20L)

try(
  add_two_numbers(10, 20)
)
```

It is also possible to add the type definition directly to the function arguments to make it visible to the user:

```{r}
multiply_two_numbers <- function(a = "integer", b = "integer:1") {
  check_args()
  a * b
}

try(
  multiply_two_numbers(10L, 1:2)
)
```

As you can see in the example above you can also check the length of the value by adding the length to the _type string_ `"integer:1"`.

### The `base_model()` function

Under the hood `typed_struct()` and `check_args()` use `base_model()` that got its name from [Pydantic's](https://docs.pydantic.dev/) `BaseModel` class. It gives you further options like validating your inputs before the assignment:

```{r}
my_model <- base_model(
  a = "integer",
  b = "double",
  .validators_before = list(
    a = as.integer,
    b = function(b) round(b, 2)
  )
)

obj <- my_model(a = 10, b = 20.123456)

typeof(obj$a)

obj
```
